### 2017-01-15 paper/howto-paper.tex
Added start of intruduce-chapter

\documentclass[10pt,conference,compsocconf]{IEEEtran}

%\usepackage{times}
%\usepackage{balance}
\usepackage{url}
\usepackage{graphicx}	% For figure environment
\usepackage{color}

\begin{document}
\title{Applying WaveNet on frequency data}

\author{
  Luzi Sennhauser\\
  Christoph Dehner\\
  Department of Computer Science, ETH Zurich, Switzerland
}

\maketitle

\begin{abstract}
  A critical part of scientific discovery is the
  communication of research findings to peers or the general public.
  Mastery of the process of scientific communication improves the
  visibility and impact of research. While this guide is a necessary
  tool for learning how to write in a manner suitable for publication
  at a scientific venue, it is by no means sufficient, on its own, to
  make its reader an accomplished writer. We also describe the rules
  for submission in the computational intelligence laboratory.
  This guide should be a
  starting point for further development of writing skills.
\end{abstract}

\section{Introduction}

In 2016 \cite{wavenet} introduced WaveNet, a generative model for raw audio achieving astonishing results in learning sound features and generating smooth audio. Trained on piano music, new highly dynamic piano music varying in melody, volume and speed smoothly could be generated.

The aim this paper is to enhance the WaveNet architecture and apply it to a more feature-oriented representation of input data.. Instead of performing a classification task on raw audio, the network is adopted to operate on the frequencies of the audio data, processed by Fouier transformation.
\textcolor{red}{Missing:\\-reference to github repo with code and results\\-Advantages/Changes by new achitecture\\-reference to the data used in the experiments}

At that
point, it is important that the reader is able to reproduce your
work~\cite{schwab00,wavelab,gentleman05}. This is why it is also
important that if the work has a computational component, the software
associated with producing the results are also made available in a
useful form. Several guidelines for making your user's experience with
your software as painless as possible is given in
Section~\ref{sec:tips-software}.

Finally, we describe the rules for submission in the computational
intelligence laboratory in Section~\ref{sec:cil}. 
This brief guide is by no means sufficient, on its own, to
make its reader an accomplished writer. The reader is urged to use the
references to further improve his or her writing skills.

\section{The Structure of a Paper}
\label{sec:structure-paper}

Scientific papers usually begin with the description of the problem,
justifying why the problem is interesting. Most importantly, it argues
that the problem is still unsolved, or that the current solutions are
unsatisfactory. This leads to the main gist of the paper, which is
``the idea''. The authors then show evidence, using derivations or
experiments, that the idea works. Since science does not occur in a
vacuum, a proper comparison to the current state of the art is often
part of the results. Following these ideas, papers usually have the
following structure:
\begin{description}
\item[Abstract] \ \\
  Short description of the whole paper, to help the
  reader decide whether to read it.
\item[Introduction] \ \\
  Describe your problem and state your
  contributions.
\item[Models and Methods] \ \\
  Describe your idea and how it was implemented to solve
  the problem. Survey the related work, giving credit where credit is
  due.
\item[Results] \ \\
  Show evidence to support your claims made in the
  introduction.
\item[Discussion] \ \\
  Discuss the strengths and weaknesses of your
  approach, based on the results. Point out the implications of your
  novel idea on the application concerned.
\item[Summary] \ \\
  Summarize your contributions in light of the new
  results.
\end{description}

\section{Tips for Good Writing}
\label{sec:tips-writing}

The ideas for good writing have come
from~\cite{editor10,jones08,anderson04}.

\subsection{Getting Help}
One should try to get a draft read by as many friendly people as
possible. And remember to treat your test readers with respect. If
they are unable to understand something in your paper, then it is
highly likely that your reviewers will not understand it
either. Therefore, do not be defensive about the criticisms you get,
but use it as an opportunity to improve the paper. Before your submit
your friends to the pain of reading your draft, please \emph{use a
  spell checker}.

\subsection{Abstract}
The abstract should really be written last, along with the title of
the paper. The four points that should be covered~\cite{jones08}:
\begin{enumerate}
\item State the problem.
\item Say why it is an interesting problem.
\item Say what your solution achieves.
\item Say what follows from your solution.
\end{enumerate}

\subsection{Figures and Tables}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{denoised_signal_1d}
  \caption{Signal compression and denoising using the Fourier basis.}
  \label{fig:denoise-fourier}
\end{figure}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\columnwidth]{local_wdenoised_1d}
  \caption{Signal compression and denoising using the Daubechies wavelet basis.}
  \label{fig:denoise-wavelet}
\end{figure}

Use examples and illustrations to clarify ideas and results. For
example, by comparing Figure~\ref{fig:denoise-fourier} and
Figure~\ref{fig:denoise-wavelet}, we can see the two different
situations where Fourier and wavelet basis perform well.

\subsection{Models and Methods}
The models and methods
section should describe what was
done to answer the research question, describe how it was done,
justify the experimental design, and
explain how the results were analyzed.

The model refers to the underlying mathematical model or structure which 
you use to describe your problem, or that your solution is based on. 
The methods on the other hand, are the algorithms used to solve the problem. 
In some cases, the suggested method directly solves the problem, without having it 
stated in terms of an underlying model. Generally though it is a better practice to have 
the model figured out and stated clearly, rather than presenting a method without specifying 
the model. In this case, the method can be more easily evaluated in the task of fitting 
the given data to the underlying model.

The methods part of this section, is not a step-by-step, directive,
protocol as you might see in your lab manual, but detailed enough such
that an interested reader can reproduce your
work~\cite{anderson04,wavelab}.

The methods section of a research paper provides the information by
which a study's validity is judged.
Therefore, it requires a clear and precise description of how an
experiment was done, and the rationale
for why specific experimental procedures were chosen.
It is usually helpful to
structure the methods section by~\cite{kallet04methods}:
\begin{enumerate}
\item Layout the model you used to describe the problem or the solution.
\item Describing the algorithms used in the study, briefly including
  details such as hyperparameter values (e.g. thresholds), and
  preprocessing steps (e.g. normalizing the data to have mean value of
  zero).
\item Explaining how the materials were prepared, for example the
  images used and their resolution.
\item Describing the research protocol, for example which examples
  were used for estimating the parameters (training) and which were
  used for computing performance.
\item Explaining how measurements were made and what
  calculations were performed. Do not reproduce the full source code in
  the paper, but explain the key steps.
\end{enumerate}

\subsection{Results}

Organize the results section based on the sequence of table and
figures you include. Prepare the tables and figures as soon as all
the data are analyzed and arrange them in the sequence that best
presents your findings in a logical way. A good strategy is to note,
on a draft of each table or figure, the one or two key results you
want to address in the text portion of the results.
The information from the figures is
summarized in Table~\ref{tab:fourier-wavelet}.

\begin{table*}[htbp]
  \centering
  \begin{tabular}[c]{|l||l|l|l|}
    \hline
    Basis&Support&Suitable signals&Unsuitable signals\\
    \hline
    Fourier&global&sine like&localized\\
    wavelet&local&localized&sine like\\
    \hline
  \end{tabular}
  \caption{Characteristics of Fourier and wavelet basis.}
  \label{tab:fourier-wavelet}
\end{table*}

When reporting computational or measurement results, always
report the mean (average value) along with a measure of variability
(standard deviation(s) or standard error of the mean).

\section{Tips for Good Software}
\label{sec:tips-software}

There is a lot of literature (for example~\cite{hunt99pragmatic} and
\cite{spolsky04software}) on how to write software. It is not the
intention of this section to replace software engineering
courses. However, in the interests of reproducible
research~\cite{schwab00}, there are a few guidelines to make your
reader happy:
\begin{itemize}
\item Have a \texttt{README} file that (at least) describes what your
  software does, and which commands to run to obtain results. Also
  mention anything special that needs to be set up, such as
  toolboxes\footnote{For those who are
  particularly interested, other common structures can be found at
  \url{http://en.wikipedia.org/wiki/README} and
  \url{http://www.gnu.org/software/womb/gnits/}.}.
\item A list of authors and contributors can be included in a file
  called \texttt{AUTHORS}, acknowledging any help that you may have
  obtained. For small projects, this information is often also
  included in the \texttt{README}.
\item Use meaningful filenames, and not \texttt{temp1.m},
  \texttt{temp2.m}. The code should also unzip into a subdirectory.
\item Document your code. Each file should at least have a short
  description about its reason for existence. Non obvious steps in the
  code should be commented.
\item Describe how the results presented in your paper can potentially
  be reproduced.
\end{itemize}

\section{Computational Intelligence Laboratory Requirements}
\label{sec:cil}

Your semester project is a group effort. It consists of four parts:
\begin{enumerate}
\item The programming assignments you solve during the semester.
\item Developing a novel solution for one of the assignments, e.g. by
  combining methods from previous programming assignments into a novel
  solution.
\item Comparing your novel solution to previous assignments.
\item Writing up your findings in a short scientific paper.
\end{enumerate}

\subsection{Developing a Novel Solution}

As your final programming assignment, you develop a novel solution to
one of the four application problems. You are free to exploit any idea
you have, provided it is not identical to any other group submission
or existing Matlab implementation of an algorithm on the
internet\footnote{\url{http://www.ethz.ch/students/semester/plagiarism_s_en.pdf}}.

Two examples for developing a novel solution:
\begin{itemize}
\item You implemented a collaborative filtering algorithm based on
  dimension reduction as part of an assignment. Now you apply
  dimension reduction to inpainting.
\item You implemented both a clustering and a sparse coding algorithm
  for image compression. Now you combine both techniques into a novel
  compression method.
\end{itemize}

\subsection{Comparison to Baselines}

You compare your novel algorithm to \emph{at least two baseline
  algorithms}. For the baselines, you can use the implementations you
developed as part of the programming assignments.

\subsection{Write Up}

The submission must be in PDF form, using the \LaTeX{} template
corresponding to the IEEE style of publication. Refer to
Section~\ref{sec:latex-primer} for more information about preparing
your document. The document should be a maximum of {\bf 4 pages}.

\subsection{\LaTeX{} Primer}
\label{sec:latex-primer}

\LaTeX{} is one of the most commonly used document preparation systems
for scientific journals and conferences. It is based on the idea
that authors should be able to focus on the content of what they are
writing without being distracted by its visual presentation.
The source of this file can be used as a starting point for how to use
the different commands in \LaTeX{}. We are using an IEEE style for
this course.

\subsubsection{Installation}

There are various different packages available for processing \LaTeX{}
documents.
On Windows, use the Mik\TeX{} package (\url{http://miktex.org/}), and
on OSX use MacTeX
(\url{http://www.tug.org/mactex/2009/}). Alternatively, on OSX, you
can install the \texttt{tetex} package via
Fink\footnote{\url{http://www.finkproject.org/}} or 
Macports\footnote{\url{http://www.macports.org/}}.

\subsubsection{Compiling \LaTeX{}}
Your directory should contain at least 4 files, in addition to image
files. Images should be in \texttt{.png}, \texttt{.jpg} or
\texttt{.pdf} format.
\begin{itemize}
\item IEEEtran.cls
\item IEEEtran.bst
\item groupXX-submission.tex
\item groupXX-literature.bib
\end{itemize}
Note that you should replace groupXX with your chosen group name.
Then, from the command line, type:
\begin{verbatim}
$ pdflatex groupXX-submission
$ bibtex groupXX-literature
$ pdflatex groupXX-submission
$ pdflatex groupXX-submission
\end{verbatim}
This should give you a PDF document \texttt{groupXX-submission.pdf}.

\subsubsection{Equations}

There are three types of equations available: inline equations, for
example $y=mx + c$, which appear in the text, unnumbered equations
$$y=mx + c,$$
which are presented on a line on its own, and numbered equations
\begin{equation}
  \label{eq:linear}
  y = mx + c
\end{equation}
which you can refer to at a later point (Equation~(\ref{eq:linear})).

\subsubsection{Tables and Figures}

Tables and figures are ``floating'' objects, which means that the text
can flow around it.
Note
that \texttt{figure*} and \texttt{table*} cause the corresponding
figure or table to span both columns.

\subsection{Grading}

There are two different types of grading criteria applied to your
project, with the corresponding weights shown in brackets.
\begin{description}
\item[Competitive] \ \\
  The following criteria is scored based on your rank
  in comparison with the rest of the class.
  \begin{itemize}
  \item time taken for computation (10\%)
  \item average rank for all other criteria relevant to the task, for
    example reconstruction error and sparsity (20\%)
  \end{itemize}
  The ranks will then be converted on a linear scale into a grade
  between 4 and 6.
\item[Non-competitive] \ \\
  The following criteria is scored based on an
  evaluation by the teaching assistants.
  \begin{itemize}
  \item quality of paper (30\%)
  \item quality of implementation (20\%)
  \item creativity of solution (20\%)
  \end{itemize}
\end{description}

\subsection{Submission System}

The deadline for submitting your project report is Friday, 22 June
2012.
You need to submit:
\begin{itemize}
\item PDF of paper.
\item Archive (\texttt{.tar.gz} or \texttt{.zip}) of software. Please
  do not forget to include author information in the source archive.
\end{itemize}

\textbf{Important:} Please check the submission instructions on the webpage 
as it is the most updated instructions.

\section{Summary}

The aim of a scientific paper is to convey the idea or discovery of
the researcher to the minds of the readers. The associated software
package provides the relevant details, which are often only briefly
explained in the paper, such that the research can be reproduced.
To write good papers, identify your key idea, make your contributions
explicit, and use examples and illustrations to describe the problems
and solutions.

\section*{Acknowledgements}
The author thanks Christian Sigg for his careful reading and helpful
suggestions.

\bibliographystyle{IEEEtran}
\bibliography{howto-paper}
\end{document}

### 2017-01-15 preprocess.py
added optional normalizing

@@ -98,7 +98,7 @@ def save_pca(input_file, output_file):
 		print("pca coefficients generation done")
 	h5f.close()
 
-def save_pca_passed_freq(freq_dict, output_file):
+def save_pca_passed_freq(freq_dict, output_file, normalize=False):
 	'''
 	Input file = frequency file
 	'''
@@ -125,23 +125,27 @@ def save_pca_passed_freq(freq_dict, output_file):
 	if VERBOSE:
 		print("pca coefficients generation done")
 
-	# calculate mean an variance
-	all_pca = np.array([]).reshape((-1,coeff_per_window))
-	for key in pca_dict:
-		all_pca = np.concatenate((all_pca, pca_dict[key]), axis=0)
-	mean = np.mean(all_pca, axis=0)
-	var = np.var(all_pca, axis=0)
+	if normalize:
+		# calculate mean an variance
+		all_pca = np.array([]).reshape((-1,coeff_per_window))
+		for key in pca_dict:
+			all_pca = np.concatenate((all_pca, pca_dict[key]), axis=0)
+		mean = np.mean(all_pca, axis=0)
+		var = np.var(all_pca, axis=0)
 
-	h5f.create_dataset('normalize/mean', data=mean)
-	h5f.create_dataset('normalize/var', data=var)
+		h5f.create_dataset('normalize/mean', data=mean)
+		h5f.create_dataset('normalize/var', data=var)
 
-	if VERBOSE:
-		print("Calculated mean and variance. Starting normalization.")
+		if VERBOSE:
+			print("Calculated mean and variance. Starting normalization.")
 
-	for key in freq_dict:
-		pca_dict[key] = pca_dict[key] - mean
-		pca_dict[key] = pca_dict[key] / (var / 10.0)
-		h5f.create_dataset('coeff/{}'.format(key), data=pca_dict[key])
+		for key in freq_dict:
+			pca_dict[key] = pca_dict[key] - mean
+			pca_dict[key] = pca_dict[key] / (var / 10.0)
+			h5f.create_dataset('coeff/{}'.format(key), data=pca_dict[key])
+	else:
+		for key in freq_dict:
+			h5f.create_dataset('coeff/{}'.format(key), data=pca_dict[key])
 
 	h5f.close()
 	print("Preprocessing done.")

### 2017-01-15 paper/howto-paper.tex
added section "model and methods"

@@
 \usepackage{color}
+\usepackage{amsmath}
 
@@
 
-At that
-point, it is important that the reader is able to reproduce your
-work~\cite{schwab00,wavelab,gentleman05}. This is why it is also
-important that if the work has a computational component, the software
-associated with producing the results are also made available in a
-useful form. Several guidelines for making your user's experience with
-your software as painless as possible is given in
-Section~\ref{sec:tips-software}.
+\section{Models and Methods}
+The following section describes the underlying model of the network used in the experiments of this paper.
+As core network the WaveNet architecture with a stack of dilated causal convolutional layers is taken.
+Hereupon, the data input layer is changed conceptually.
+\subsection{Data preprocessing}
+To extract frequencies, the raw audio data consisting of pressure measurements are divided into chunks of size 2048, which are then transformed into the frequency space using FFT.
+Therefrom, the first 150 coefficients are taken as dominant frequencies for the subsequent training.
+To ensure that audio features are transformed smoothly, two neighboring raw audio chunks overlap in a few samples, described by a variable $stride$.
+Equation \ref{eq:fft} summarized the frequency extraction of the $i^\textnormal{th}$ chunk compactly.
+\begin{multline}\label{eq:fft}
+\textnormal{frequencies\_coeffs}_i =
+FFT(\textnormal{inputData}[i \cdot stride:\\i \cdot stride+2048])[:150]
+\end{multline}
 
-Finally, we describe the rules for submission in the computational
-intelligence laboratory in Section~\ref{sec:cil}. 
-This brief guide is by no means sufficient, on its own, to
-make its reader an accomplished writer. The reader is urged to use the
-references to further improve his or her writing skills.
+Considering real and imaginary parts of the complex Fourier coefficients separately, each chunk of 2048 raw audio samples is transformed into 300 real frequency values.
+
+To decrease the dimensionality even further, principle component analysis is used to reduce these 300 frequencies to 100 values.
+These 100 coefficient are considered one tone and fed into the network as one time step.
+Whereas in the original WaveNet architecture, 16000 samples represented on second of audio, the adopted network only needs to process $16000/2048 \approx 8$ samples per second.
+However, this change infers a fundamentally different approach of audio training and generation, described in the next section.
+
+\textcolor{red}{Missing: Comparison/error plot of original data vs. preprocessed data}
+
+\subsection{Network architecture and loss}
+Whereas the original WaveNet paper designed the network to perform a classification task, the adopted network aims to train and predict the frequencies of a audio sample directly.
+The network input during one time step is now not only one class describing the current tone, but a 100 dimensional vector representing the frequencies of the tone of time step in a compressed way.
+Similarly, the output of the network is no longer a softmax layer but also a 100 dimensional vector.
+
+This change in architecture also affects the loss during training. Instead of a softmax cross-entropy loss the adopted network uses the mean squared error as loss function during training.
 

### 2017-01-16 paper/howto-paper.tex
extended section "introduction"

@@
 
-The aim this paper is to enhance the WaveNet architecture and apply it to a more feature-oriented representation of input data..
+The aim this paper is to enhance the WaveNet architecture and apply it to a more feature-oriented representation of input data.
 Instead of performing a classification task on raw audio, the network is adopted to operate on the frequencies of the audio data, processed by Fouier transformation.
+In particular, the following topics are covered:
+\begin{itemize}
+\item Based on WaveNet from \cite{wavent} a network architecture to generate music is introduced operating on the frequencies of audio input.
+\item This new network is trained and evaluated on piano music from the MusicNet dataset.
+Furthermore, optimization approaches as data normalization and variations in preprocessing the data are included in the experiments.
+\item Results from the evaluation are analyzed in detail. In the context of the introduced network, advantages as well as weaknesses are discussed. 
+\end{itemize}
 \textcolor{red}{Missing:\\-reference to github repo with code and results\\-Advantages/Changes by new achitecture\\-reference to the data used in the experiments}
@@
 Hereupon, the data input layer is changed conceptually.
+\subsection{WaveNet architecture}
+\textcolor{red}{rough description of wavenet architecture and specification used during the experiments}
 \subsection{Data preprocessing}

### 2017-01-16 paper/howto-paper.tex
added text to introduction

@@
 \section{Introduction}
-
-In 2016 \cite{wavenet} introduced WaveNet, a generative model for raw audio achieving astonishing results in learning sound features and generating smooth audio.
-Trained on piano music, new highly dynamic piano music varying in melody, volume and speed smoothly could be generated.
-
+In 2016, \cite{wavenet} introduced WaveNet, a generative model for raw audio achieving astonishing results in learning sound features and generating smooth audio.
+Trained on piano music, new highly dynamic piano music varying in melody, volume and speed smoothly could be generated.\\
 The aim this paper is to enhance the WaveNet architecture and apply it to a more feature-oriented representation of input data.
@@
 \begin{itemize}
-\item Based on WaveNet from \cite{wavent} a network architecture to generate music is introduced operating on the frequencies of audio input.
+\item Based on WaveNet from \cite{wavenet} a network architecture to generate music is introduced operating on the frequencies of audio input.
 \item This new network is trained and evaluated on piano music from the MusicNet dataset.
@@
 \end{itemize}
-\textcolor{red}{Missing:\\-reference to github repo with code and results\\-Advantages/Changes by new achitecture\\-reference to the data used in the experiments}
+As training data in the experiments, piano samples from the MusicNet dataset, a collection of 330 freely-licenced recordings of classical music are taken.
+The data is publicly available online \cite{musicnet}.\\
+The source code of the experiments can be found on GitHub.\footnote{\url{https://github.com/sennluzi/tensorflow-wavenet}}
 
@@
 \end{multline}
-
-Considering real and imaginary parts of the complex Fourier coefficients separately, each chunk of 2048 raw audio samples is transformed into 300 real frequency values.
-
+Considering real and imaginary parts of the complex Fourier coefficients separately, each chunk of 2048 raw audio samples is transformed into 300 real frequency values.\\
 To decrease the dimensionality even further, principle component analysis is used to reduce these 300 frequencies to 100 values.
@@
 The network input during one time step is now not only one class describing the current tone, but a 100 dimensional vector representing the frequencies of the tone of time step in a compressed way.
-Similarly, the output of the network is no longer a softmax layer but also a 100 dimensional vector.
-
+Similarly, the output of the network is no longer a softmax layer but also a 100 dimensional vector.\\
 This change in architecture also affects the loss during training. Instead of a softmax cross-entropy loss the adopted network uses the mean squared error as loss function during training.

### 2017-01-18 paper/paper.tex
added figures to paper
completed WaveNet section in paper and added one more figure

--- 2017-01-16 paper/howto-paper.tex
@@
 The following section describes the underlying model of the network used in the experiments of this paper.
-As core network the WaveNet architecture with a stack of dilated causal convolutional layers is taken.
+As core network the WaveNet architecture with a stack of dilated causal convolutional layers (figure \ref{fig:wavenet_dilated_cnn}) is taken.
 Hereupon, the data input layer is changed conceptually.
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/wavenet_dilated_cnn.png}
+  \caption{dilated convolutional network}
+  \label{fig:wavenet_dilated_cnn}
+\end{figure}
+
 \subsection{WaveNet architecture}
+Given a sequence of amplitudes, WaveNet predicts the subsequent amplitude.
+The amplitudes are quantized in 256 bins \cite{itu1988711}.
+In the $i$th step, the network then outputs a probability distribution $P(x_{i+1} | x_1, ..., x_i)$.
+The amplitude $x_{i+1}$ of the next step is then sampled according to that distribution.\\
+In WaveNet, training can be done in parallel while the generation of a audio sample is done sequentially.\\
+As a basis for WaveNet, the implementation of Igor Babuschkin is taken.\footnote{\url{https://github.com/ibab/tensorflow-wavenet}}\\
+The original WaveNet paper mentiones only very few coefficients for how their network is configured.
+Therefore reproducing the paper exactly is very hard.
+The following numbers are assumptions mainly based on the experience of Babuschkin's implementation.
+In the following experiments, the dilations are $2^n$ for $n$ taking the values $(0,1,...,7,8)$ and are repeated 5 times.
+This leads to 50 hidden dilation layers.
+The convolution filter size is set to $2$ and the number of channels of the residual and the dilation is $32$.\\
+The input of the WaveNet network is a sequence of amplitudes, each of which encoded as a one-hot vector.
+
 \textcolor{red}{rough description of wavenet architecture and specification used during the experiments}
@@
 Considering real and imaginary parts of the complex Fourier coefficients separately, each chunk of 2048 raw audio samples is transformed into 300 real frequency values.\\
-To decrease the dimensionality even further, principle component analysis is used to reduce these 300 frequencies to 100 values.
-These 100 coefficient are considered one tone and fed into the network as one time step.
+To decrease the dimensionality even further, principle component analysis (PCA) is used to reduce these 300 frequencies to 100 values.
+These 100 PCA coefficient are considered one tone and fed into the network as one time step.\\
+The eigenvectors of the PCA are shown in figure \ref{fig:pca_eigenvectors_imaginary_numbers}.
+Figure \ref{fig:pca_eigenvectors} illustrates how the PCA looks like when taking the absolute values of the imaginary numbers, out of which the components are built from.
+In this figure, one can see, that the relevant frequencies are all relatively low.
+Figure \ref{fig:pca_explained_variance} shows that a vast majority of the variance of the frequencies is captured by the first 100 PCA components.
+Summarizing, in neither of the two steps, we loose much of the original input information.
+This should lead to a decent compression of the audio signal.\\
+To make this evident, one can figure invert all the steps and plot the compressed wave and the original wave (see \ref{fig:preprocessing_difference_waveform}).
+It is easy to see that they resemble a lot.
+After preprocessing the audio contains a cracking noise, but nevertheless is the melody and the tone of the music well audible.\\
+To ensure this preprocessing still preserves the nature of the music, one can invert all steps.
+Although after preprocessing, there is an audible cracking, the music then is still clearly hearable and the pitches recognizable.
+On one hand, the waveform looks still similar (figure \ref{fig:preprocessing_difference_waveform})
 Whereas in the original WaveNet architecture, 16000 samples represented on second of audio, the adopted network only needs to process $16000/2048 \approx 8$ samples per second.
@@
 
-\textcolor{red}{Missing: Comparison/error plot of original data vs. preprocessed data}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/pca_eigenvectors_imaginary_numbers.png}
+  \caption{Eigenvectors of the PCA where the imaginary and the real part of the FFT are stored as a separate value in the PCA component.}
+  \label{fig:pca_eigenvectors_imaginary_numbers}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/pca_eigenvectors.png}
+  \caption{Transformed eigenvectors of the PCA: The values of the vectors are the absolute value of the corresponding imaginary number}
+  \label{fig:pca_eigenvectors}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/pca_explained_variance}
+  \caption{Signal compression and denoising using the Daubechies wavelet basis.}
+  \label{fig:pca_explained_variance}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/preprocessing_difference_waveform}
+  \caption{An example of how the output audio differs before and after preprocessing (compression)}
+  \label{fig:preprocessing_difference_waveform}
+\end{figure}
 
@@
 \subsection{Figures and Tables}
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\columnwidth]{denoised_signal_1d}
-  \caption{Signal compression and denoising using the Fourier basis.}
-  \label{fig:denoise-fourier}
-\end{figure}
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\columnwidth]{local_wdenoised_1d}
-  \caption{Signal compression and denoising using the Daubechies wavelet basis.}
-  \label{fig:denoise-wavelet}
-\end{figure}
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{howto-paper}
+\bibliography{paper}
 \end{document}

### 2017-01-18 paper/paper.tex
started with result and discussion section

@@
 In WaveNet, training can be done in parallel while the generation of a audio sample is done sequentially.\\
-As a basis for WaveNet, the implementation of Igor Babuschkin is taken.\footnote{\url{https://github.com/ibab/tensorflow-wavenet}}\\
+As a basis for WaveNet, the implementation of Igor Babuschkin is taken \cite{Babuschkin2016}.\\
 The original WaveNet paper mentiones only very few coefficients for how their network is configured.
@@
 
-\section{The Structure of a Paper}
-\label{sec:structure-paper}
+\section{Results}
+The generated audio files start with some fastly changing frequencies and then after 30-40 frames (1-2 seconds) it stabilizes on one distribution of frequencies.
+From then on the values of the frequencies do not change any more.
+Figures \ref{fig:original_frequencies} and \ref{fig:generated_frequencies} show that the neural network is not able to produce a change of frequencies while a real piece of music observes a change every 20-200 frames.\\
+The consequence of these frequencies not changing is, that the audio file consists after a noise at the beginning only of a steady pitch. And it doesn't resemble a pitch either.\\
 
-Scientific papers usually begin with the description of the problem,
-justifying why the problem is interesting. Most importantly, it argues
-that the problem is still unsolved, or that the current solutions are
-unsatisfactory. This leads to the main gist of the paper, which is
-``the idea''. The authors then show evidence, using derivations or
-experiments, that the idea works. Since science does not occur in a
-vacuum, a proper comparison to the current state of the art is often
-part of the results. Following these ideas, papers usually have the
-following structure:
-\begin{description}
-\item[Abstract] \ \\
-  Short description of the whole paper, to help the
-  reader decide whether to read it.
-\item[Introduction] \ \\
-  Describe your problem and state your
-  contributions.
-\item[Models and Methods] \ \\
-  Describe your idea and how it was implemented to solve
-  the problem. Survey the related work, giving credit where credit is
-  due.
-\item[Results] \ \\
-  Show evidence to support your claims made in the
-  introduction.
-\item[Discussion] \ \\
-  Discuss the strengths and weaknesses of your
-  approach, based on the results. Point out the implications of your
-  novel idea on the application concerned.
-\item[Summary] \ \\
-  Summarize your contributions in light of the new
-  results.
-\end{description}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/original_frequencies.png}
+  \caption{Frequencies of a piano piece used for training}
+  \label{fig:original_frequencies}
+\end{figure}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/generated_frequencies_20s.png}
+  \caption{Frequencies of the generated audio piece}
+  \label{fig:generated_frequencies}
+\end{figure}
 
-\section{Tips for Good Writing}
-\label{sec:tips-writing}
+\begin{figure*}[tbp]
+  \centering
+  \includegraphics[width=\textwidth]{figures/loss.png}
+  \caption{Training error}
+  \label{fig:training_error}
+\end{figure*}
 
-The ideas for good writing have come
-from~\cite{editor10,jones08,anderson04}.
-
-\subsection{Getting Help}
-One should try to get a draft read by as many friendly people as
-possible. And remember to treat your test readers with respect. If
-they are unable to understand something in your paper, then it is
-highly likely that your reviewers will not understand it
-either. Therefore, do not be defensive about the criticisms you get,
-but use it as an opportunity to improve the paper. Before your submit
-your friends to the pain of reading your draft, please \emph{use a
-  spell checker}.
-
-\subsection{Abstract}
-The abstract should really be written last, along with the title of
-the paper. The four points that should be covered~\cite{jones08}:
-\begin{enumerate}
-\item State the problem.
-\item Say why it is an interesting problem.
-\item Say what your solution achieves.
-\item Say what follows from your solution.
-\end{enumerate}
-
-\subsection{Figures and Tables}
-
-Use examples and illustrations to clarify ideas and results. For
-example, by comparing Figure~\ref{fig:denoise-fourier} and
-Figure~\ref{fig:denoise-wavelet}, we can see the two different
-situations where Fourier and wavelet basis perform well.
-
-\subsection{Models and Methods}
-The models and methods
-section should describe what was
-done to answer the research question, describe how it was done,
-justify the experimental design, and
-explain how the results were analyzed.
-
-The model refers to the underlying mathematical model or structure which 
-you use to describe your problem, or that your solution is based on. 
-The methods on the other hand, are the algorithms used to solve the problem. 
-In some cases, the suggested method directly solves the problem, without having it 
-stated in terms of an underlying model. Generally though it is a better practice to have 
-the model figured out and stated clearly, rather than presenting a method without specifying 
-the model. In this case, the method can be more easily evaluated in the task of fitting 
-the given data to the underlying model.
-
-The methods part of this section, is not a step-by-step, directive,
-protocol as you might see in your lab manual, but detailed enough such
-that an interested reader can reproduce your
-work~\cite{anderson04,wavelab}.
-
-The methods section of a research paper provides the information by
-which a study's validity is judged.
-Therefore, it requires a clear and precise description of how an
-experiment was done, and the rationale
-for why specific experimental procedures were chosen.
-It is usually helpful to
-structure the methods section by~\cite{kallet04methods}:
-\begin{enumerate}
-\item Layout the model you used to describe the problem or the solution.
-\item Describing the algorithms used in the study, briefly including
-  details such as hyperparameter values (e.g. thresholds), and
-  preprocessing steps (e.g. normalizing the data to have mean value of
-  zero).
-\item Explaining how the materials were prepared, for example the
-  images used and their resolution.
-\item Describing the research protocol, for example which examples
-  were used for estimating the parameters (training) and which were
-  used for computing performance.
-\item Explaining how measurements were made and what
-  calculations were performed. Do not reproduce the full source code in
-  the paper, but explain the key steps.
-\end{enumerate}
-
-\subsection{Results}
-
-Organize the results section based on the sequence of table and
-figures you include. Prepare the tables and figures as soon as all
-the data are analyzed and arrange them in the sequence that best
-presents your findings in a logical way. A good strategy is to note,
-on a draft of each table or figure, the one or two key results you
-want to address in the text portion of the results.
-The information from the figures is
-summarized in Table~\ref{tab:fourier-wavelet}.
-
-\begin{table*}[htbp]
-  \centering
-  \begin{tabular}[c]{|l||l|l|l|}
-    \hline
-    Basis&Support&Suitable signals&Unsuitable signals\\
-    \hline
-    Fourier&global&sine like&localized\\
-    wavelet&local&localized&sine like\\
-    \hline
-  \end{tabular}
-  \caption{Characteristics of Fourier and wavelet basis.}
-  \label{tab:fourier-wavelet}
-\end{table*}
-
-When reporting computational or measurement results, always
-report the mean (average value) along with a measure of variability
-(standard deviation(s) or standard error of the mean).
-
-\section{Tips for Good Software}
-\label{sec:tips-software}
-
-There is a lot of literature (for example~\cite{hunt99pragmatic} and
-\cite{spolsky04software}) on how to write software. It is not the
-intention of this section to replace software engineering
-courses. However, in the interests of reproducible
-research~\cite{schwab00}, there are a few guidelines to make your
-reader happy:
-\begin{itemize}
-\item Have a \texttt{README} file that (at least) describes what your
-  software does, and which commands to run to obtain results. Also
-  mention anything special that needs to be set up, such as
-  toolboxes\footnote{For those who are
-  particularly interested, other common structures can be found at
-  \url{http://en.wikipedia.org/wiki/README} and
-  \url{http://www.gnu.org/software/womb/gnits/}.}.
-\item A list of authors and contributors can be included in a file
-  called \texttt{AUTHORS}, acknowledging any help that you may have
-  obtained. For small projects, this information is often also
-  included in the \texttt{README}.
-\item Use meaningful filenames, and not \texttt{temp1.m},
-  \texttt{temp2.m}. The code should also unzip into a subdirectory.
-\item Document your code. Each file should at least have a short
-  description about its reason for existence. Non obvious steps in the
-  code should be commented.
-\item Describe how the results presented in your paper can potentially
-  be reproduced.
-\end{itemize}
-
-\section{Computational Intelligence Laboratory Requirements}
-\label{sec:cil}
-
-Your semester project is a group effort. It consists of four parts:
-\begin{enumerate}
-\item The programming assignments you solve during the semester.
-\item Developing a novel solution for one of the assignments, e.g. by
-  combining methods from previous programming assignments into a novel
-  solution.
-\item Comparing your novel solution to previous assignments.
-\item Writing up your findings in a short scientific paper.
-\end{enumerate}
-
-\subsection{Developing a Novel Solution}
-
-As your final programming assignment, you develop a novel solution to
-one of the four application problems. You are free to exploit any idea
-you have, provided it is not identical to any other group submission
-or existing Matlab implementation of an algorithm on the
-internet\footnote{\url{http://www.ethz.ch/students/semester/plagiarism_s_en.pdf}}.
-
-Two examples for developing a novel solution:
-\begin{itemize}
-\item You implemented a collaborative filtering algorithm based on
-  dimension reduction as part of an assignment. Now you apply
-  dimension reduction to inpainting.
-\item You implemented both a clustering and a sparse coding algorithm
-  for image compression. Now you combine both techniques into a novel
-  compression method.
-\end{itemize}
-
-\subsection{Comparison to Baselines}
-
-You compare your novel algorithm to \emph{at least two baseline
-  algorithms}. For the baselines, you can use the implementations you
-developed as part of the programming assignments.
-
-\subsection{Write Up}
-
-The submission must be in PDF form, using the \LaTeX{} template
-corresponding to the IEEE style of publication. Refer to
-Section~\ref{sec:latex-primer} for more information about preparing
-your document. The document should be a maximum of {\bf 4 pages}.
-
-\subsection{\LaTeX{} Primer}
-\label{sec:latex-primer}
-
-\LaTeX{} is one of the most commonly used document preparation systems
-for scientific journals and conferences. It is based on the idea
-that authors should be able to focus on the content of what they are
-writing without being distracted by its visual presentation.
-The source of this file can be used as a starting point for how to use
-the different commands in \LaTeX{}. We are using an IEEE style for
-this course.
-
-\subsubsection{Installation}
-
-There are various different packages available for processing \LaTeX{}
-documents.
-On Windows, use the Mik\TeX{} package (\url{http://miktex.org/}), and
-on OSX use MacTeX
-(\url{http://www.tug.org/mactex/2009/}). Alternatively, on OSX, you
-can install the \texttt{tetex} package via
-Fink\footnote{\url{http://www.finkproject.org/}} or 
-Macports\footnote{\url{http://www.macports.org/}}.
-
-\subsubsection{Compiling \LaTeX{}}
-Your directory should contain at least 4 files, in addition to image
-files. Images should be in \texttt{.png}, \texttt{.jpg} or
-\texttt{.pdf} format.
-\begin{itemize}
-\item IEEEtran.cls
-\item IEEEtran.bst
-\item groupXX-submission.tex
-\item groupXX-literature.bib
-\end{itemize}
-Note that you should replace groupXX with your chosen group name.
-Then, from the command line, type:
-\begin{verbatim}
-$ pdflatex groupXX-submission
-$ bibtex groupXX-literature
-$ pdflatex groupXX-submission
-$ pdflatex groupXX-submission
-\end{verbatim}
-This should give you a PDF document \texttt{groupXX-submission.pdf}.
-
-\subsubsection{Equations}
-
-There are three types of equations available: inline equations, for
-example $y=mx + c$, which appear in the text, unnumbered equations
-$$y=mx + c,$$
-which are presented on a line on its own, and numbered equations
-\begin{equation}
-  \label{eq:linear}
-  y = mx + c
-\end{equation}
-which you can refer to at a later point (Equation~(\ref{eq:linear})).
-
-\subsubsection{Tables and Figures}
-
-Tables and figures are ``floating'' objects, which means that the text
-can flow around it.
-Note
-that \texttt{figure*} and \texttt{table*} cause the corresponding
-figure or table to span both columns.
-
-\subsection{Grading}
-
-There are two different types of grading criteria applied to your
-project, with the corresponding weights shown in brackets.
-\begin{description}
-\item[Competitive] \ \\
-  The following criteria is scored based on your rank
-  in comparison with the rest of the class.
-  \begin{itemize}
-  \item time taken for computation (10\%)
-  \item average rank for all other criteria relevant to the task, for
-    example reconstruction error and sparsity (20\%)
-  \end{itemize}
-  The ranks will then be converted on a linear scale into a grade
-  between 4 and 6.
-\item[Non-competitive] \ \\
-  The following criteria is scored based on an
-  evaluation by the teaching assistants.
-  \begin{itemize}
-  \item quality of paper (30\%)
-  \item quality of implementation (20\%)
-  \item creativity of solution (20\%)
-  \end{itemize}
-\end{description}
-
-\subsection{Submission System}
-
-The deadline for submitting your project report is Friday, 22 June
-2012.
-You need to submit:
-\begin{itemize}
-\item PDF of paper.
-\item Archive (\texttt{.tar.gz} or \texttt{.zip}) of software. Please
-  do not forget to include author information in the source archive.
-\end{itemize}
-
-\textbf{Important:} Please check the submission instructions on the webpage 
-as it is the most updated instructions.
+\section{Discussion}
+The approach of this paper could be promising due to the following reason: The original WaveNet paper's network has to learn the behaviour of sound consisting of repetitive oscillations.
+With transforming the signal to the Fourier space, we take this burden from the neural network.
+The network can focus on data which is more related to the final tone.
+For example when a tone is repeated for 1s, this is also the same output of the network.\\
+But it might be exactly this phenomenon which makes good music generation impossible.
+During training, the network might do really well only staying on one and the same pitch, because changes happen in too rare occasions to influence the overall loss.\\
+When improving this approach further, one could try to include the rate of changing tones into the loss function and to punish a network when producing too many times in a row a similar output.\\
 
 \section{Summary}
-
-The aim of a scientific paper is to convey the idea or discovery of
-the researcher to the minds of the readers. The associated software
-package provides the relevant details, which are often only briefly
-explained in the paper, such that the research can be reproduced.
-To write good papers, identify your key idea, make your contributions
-explicit, and use examples and illustrations to describe the problems
-and solutions.
-
-\section*{Acknowledgements}
-The author thanks Christian Sigg for his careful reading and helpful
-suggestions.
 

### 2017-01-19 paper/paper.tex
completed discussion part of the paper and added a first draft of the summary

@@
 The network can focus on data which is more related to the final tone.
-For example when a tone is repeated for 1s, this is also the same output of the network.\\
+For example while a tone is the same for a certain time, the output of the network is also the same.\\
 But it might be exactly this phenomenon which makes good music generation impossible.
 During training, the network might do really well only staying on one and the same pitch, because changes happen in too rare occasions to influence the overall loss.\\
-When improving this approach further, one could try to include the rate of changing tones into the loss function and to punish a network when producing too many times in a row a similar output.\\
+Future work based on that paper, one could try to include the rate of changing tones into the loss function and to punish a network when producing too many times in a row a similar output.\\
+There are two reasons one could think of, why this approach's result does not match its expectations:
+\begin{itemize}
+  \item The original paper quantizes the amplitudes and can then use a softmax layer as a last layer of the network.
+In this new approach, we need continous values (PCA coefficients).
+Since the results are much worse than with WaveNet, the missing quantization (and connected to it the missing softmax layer) might be the reason behind this phenomenon.
+  \item The original WaveNet has a probability distribution as output.
+The exact amplitudes are then sampled according to that distribution.
+In the approach presented in this paper, one relies on continuous values, which are directly calculated (no probabilities included).
+In future work, one can change the model to produce a probability distribution of the output values.
+\end{itemize}
 
 \section{Summary}
+With WaveNet an already very promising music generation network is presented.
+It relies on predicting amplitudes when the preceeding amplitudes are given.
+In this paper it is investigated whether transforming the data to the Fourier space improves the predictions.
+After transforming, the dimensionality of the audio file is vastly diminished.
+But at the same time, the ability to quantize the network's outputs is lost.
+Unfortunately the network returned results not as good as with the original WaveNet paper.
+With the new method, the generated audio files rapidly converge to one steady tone.
+The changing behaviour of pitches is not learned by the network.
 

### 2017-01-20 paper/paper.tex
Revied section "Introduction" and "Models and Methods"

@@
 \section{Introduction}
-In 2016, \cite{wavenet} introduced WaveNet, a generative model for raw audio achieving astonishing results in learning sound features and generating smooth audio.
-Trained on piano music, new highly dynamic piano music varying in melody, volume and speed smoothly could be generated.\\
-The aim this paper is to enhance the WaveNet architecture and apply it to a more feature-oriented representation of input data.
-Instead of performing a classification task on raw audio, the network is adopted to operate on the frequencies of the audio data, processed by Fouier transformation.
+In 2016, \cite{wavenet} introduced WaveNet, a generative model for raw audio achieving astonishing results in learning sound features and generating harmonious audio.
+Trained on piano music, highly dynamic piano music varying smoothly in melody, volume and speed could be generated.\\
+The aim of this paper is to enhance the WaveNet architecture and apply it to a more feature-oriented representation of input data.
+Instead of performing a classification task on raw audio, the network is adopted to operate on the frequencies of the audio data, processed by Fouier transformation.\\
+As training data in the experiments, piano samples from the MusicNet dataset, a collection of 330 freely-licenced recordings of classical music are taken.
+The data is publicly available online \cite{musicnet}.\\
 In particular, the following topics are covered:
 \begin{itemize}
-\item Based on WaveNet from \cite{wavenet} a network architecture to generate music is introduced operating on the frequencies of audio input.
+\item Based on WaveNet from \cite{wavenet}, a network architecture to generate music is introduced operating on the frequencies of audio input.
 \item This new network is trained and evaluated on piano music from the MusicNet dataset.
@@
 \end{itemize}
-As training data in the experiments, piano samples from the MusicNet dataset, a collection of 330 freely-licenced recordings of classical music are taken.
-The data is publicly available online \cite{musicnet}.\\
 The source code of the experiments can be found on GitHub.\footnote{\url{https://github.com/sennluzi/tensorflow-wavenet}}
@@
 The following section describes the underlying model of the network used in the experiments of this paper.
-As core network the WaveNet architecture with a stack of dilated causal convolutional layers (figure \ref{fig:wavenet_dilated_cnn}) is taken.
+As core network the WaveNet architecture with a stack of dilated causal convolutional layers is taken.
 Hereupon, the data input layer is changed conceptually.
@@
 \end{figure}
-
 \subsection{WaveNet architecture}
 Given a sequence of amplitudes, WaveNet predicts the subsequent amplitude.
-The amplitudes are quantized in 256 bins \cite{itu1988711}.
-In the $i$th step, the network then outputs a probability distribution $P(x_{i+1} | x_1, ..., x_i)$.
+This task is done in a classification setting: The amplitudes are quantized in 256 classes \cite{itu1988711}.
+The input of the WaveNet network is a sequence of such quantizied amplitudes, each encoded as a 256-dimensional one-hot vector.
+Performing a softmax-function in its last layer, the network outputs a probability distribution $P(x_{i+1} | x_1, ..., x_i)$ in the $i$th step.
 The amplitude $x_{i+1}$ of the next step is then sampled according to that distribution.\\
-In WaveNet, training can be done in parallel while the generation of a audio sample is done sequentially.\\
-As a basis for WaveNet, the implementation of Igor Babuschkin is taken \cite{Babuschkin2016}.\\
-The original WaveNet paper mentiones only very few coefficients for how their network is configured.
+To model the the dependency of an audio sample on previous samples, WaveNet uses a stack of dilated convolutions.
+Figure \ref{fig:wavenet_dilated_cnn} depicts this architecture exemplarily for such a network of depth 4.
+By using exponentially growing dilations in depth, a network with $n$ dilated convolutional layers can observe the last $2^n$ input values to predict an output.\\
+Training WaveNet is done similar as an autoencoder: The $i$th amplitude of the input data is predicted by only using previous input data and taken as label for the loss calculation and backpropagation.
+Thus, training only requires the input data and can be done in parallel.
+Against that, the generation of a audio sample has to be processed sequentially, since the $i$th step relies on the output of the $(i-1)$th one.\\
+As a basis for WaveNet, the implementation of Igor Babuschkin is taken \cite{Babuschkin2016}.
+The original WaveNet paper mentions only very few coefficients for how their network is configured.
 Therefore reproducing the paper exactly is very hard.
 The following numbers are assumptions mainly based on the experience of Babuschkin's implementation.
-In the following experiments, the dilations are $2^n$ for $n$ taking the values $(0,1,...,7,8)$ and are repeated 5 times.
+In the following experiments, the dilations are $2^n$ for $n$ taking the values $(0,1,...,8)$ and are repeated 5 times.
 This leads to 50 hidden dilation layers.
 The convolution filter size is set to $2$ and the number of channels of the residual and the dilation is $32$.\\
-The input of the WaveNet network is a sequence of amplitudes, each of which encoded as a one-hot vector.
-
-\textcolor{red}{rough description of wavenet architecture and specification used during the experiments}
 \subsection{Data preprocessing}

### 2017-01-20 paper/paper.tex
Added section "results"

@@
 
+\pagestyle{plain}
 \begin{document}
@@
 \section{Results}
-The generated audio files start with some fastly changing frequencies and then after 30-40 frames (1-2 seconds) it stabilizes on one distribution of frequencies.
-From then on the values of the frequencies do not change any more.
-Figures \ref{fig:original_frequencies} and \ref{fig:generated_frequencies} show that the neural network is not able to produce a change of frequencies while a real piece of music observes a change every 20-200 frames.\\
-The consequence of these frequencies not changing is, that the audio file consists after a noise at the beginning only of a steady pitch. And it doesn't resemble a pitch either.\\
-
+The results of all experiments are summarized in a python notebook of the GitHub repository.\footnote{"evaluate\_results.ipynb", available in the subfolder "notebooks"} Unfortunately, the adopted network could not perform as well as the original WaveNet architecture.
+All generated audio files stabilized on one distribution of frequencies after a few seconds.
+From then on the values of the frequencies did not change any more.
+Instead of learning the complete range of frequencies, which define the features of the audio input, the network could only learn some low basis frequencies being constantly present all the time.\\
+Figures \ref{fig:original_frequencies} and \ref{fig:generated_frequencies} compare the frequencies of a recorded audio piece from the input data and the frequencies of a generated sample.
+In the first 200 time frames, the frequency distributions look quite similar.
+However, afterwards the neural network is not able to produce frequency changes continuously, while a real piece of music observes a change every 20-60 frames.\\
+Furthermore the original and generated audios differ in their frequency range.
+Whereas the original music piece only consists of frequencies up to 32, the generated audio contains frequencies up to 105.
+This shows that the network has difficulties to decide which frequencies are important and cannot produce as compact and compressed audio output as it is achieved by using FFT and PCA on recorded audio data.\\
+Nevertheless, the experiments of this paper can be analysed more detailled. Varying the stride and including normalization yielded conceptually different results.\\
+Overlapping chunks of input data turned out to be essential for learning audio features properly.
+With a stride of 512 (resulting in an overlap of 75\% of neighboring data chunks) results with dynamic volume changes and frequencies within the range $[0, 105]$ within the first seconds of the generated audio could be achieved.
+In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.\textcolor{red}{Missing for discussion: on the one hand overlap necessary, on the other hand emphasises this to learn constant tones.}\\
+A second conceptual attempt during the experiments was to normalize the PCA results before the training.
+The aim of this additional preprocessing task was to balance the contribution of all 100 PCA dimensions to the loss, hoping that all frequencies are learned more equally.
+In a first attempt, data normalzed to zero mean and unit variance turned out to produce too small loss for training.
+With vanishing gradients the network could not learn at all.
+Therefore, in a second attempt the input data was normalized to zero mean and an equal variance of all dimensions of 10.
+With this specifications, the network was able to learn as well as without normalization.\\
+As intended, the normalization helped to learn higher frequencies.
+The generated results with this setting contained more different frequencies from the range $[0,360]$.
+But the generated audio still converged to a constant distribution of frequencies after a few seconds.
 \begin{figure}[tbp]
@@
   \centering
-  \includegraphics[width=\columnwidth]{figures/generated_frequencies_20s.png}
+  \includegraphics[width=\columnwidth]{figures/generated_frequencies_20s_2.png}
   \caption{Frequencies of the generated audio piece}

### 2017-01-20 paper/paper.tex
added figure comments to "results".
commented out training error plot

@@
 With a stride of 512 (resulting in an overlap of 75\% of neighboring data chunks) results with dynamic volume changes and frequencies within the range $[0, 105]$ within the first seconds of the generated audio could be achieved.
+Figure \ref{fig:generated_frequencies} visualizes the frequencies of the first 30 seconds of the output.
 In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.\textcolor{red}{Missing for discussion: on the one hand overlap necessary, on the other hand emphasises this to learn constant tones.}\\
@@
   \includegraphics[width=\columnwidth]{figures/original_frequencies.png}
-  \caption{Frequencies of a piano piece used for training}
+  \caption{Frequencies of a piano piece used for training (Recording 2322 from the MusicNet dataset). }
   \label{fig:original_frequencies}
@@
   \includegraphics[width=\columnwidth]{figures/generated_frequencies_20s_2.png}
-  \caption{Frequencies of the generated audio piece}
+  \caption{Frequencies of a generated audio piece, model trained on the MusicNet dataset with $stride=512$ and no normalization.}
   \label{fig:generated_frequencies}
@@
 
-\begin{figure*}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{figures/loss.png}
-  \caption{Training error}
-  \label{fig:training_error}
-\end{figure*}
+%\begin{figure*}[tbp]
+%  \centering
+%  \includegraphics[width=\textwidth]{figures/loss.png}
+%  \caption{Training error}
+%  \label{fig:training_error}
+%\end{figure*}
 

### 2017-01-20 generate.py
code commented and readme written

@@ -75,7 +75,7 @@ def get_arguments():
         '--pca_out_path',
         type=str,
         default=None,
-        help='Path to output wav file')
+        help='Path to output pca file')
     parser.add_argument(
         '--save_every',
         type=int,
@@ -105,6 +105,10 @@ def write_wav(waveform, sample_rate, filename):
     librosa.output.write_wav(filename, y, sample_rate)
     print('Updated wav file at {}'.format(filename))
 
+'''
+EXTENSION
+This function writes the pca coefficients to a file (instead of previously directly the audio file)
+'''
 def write_pca(waveform, sample_rate, filename):
     y = np.array(waveform)
     h5f = h5py.File(filename, 'w')
@@ -117,9 +121,10 @@ def create_seed(filename,
                 quantization_channels,
                 window_size=WINDOW,
                 silence_threshold=SILENCE_THRESHOLD):
-    #audio, _ = librosa.load('/mnt/c/Users/Luzi/Music/williamson-a_few_things_to_hear_before_we_all_blow_up-12-a_please_goodbye_from_whore-59-88.mp3.wav', sr=sample_rate, mono=True)
-    #audio = audio_reader.trim_silence(audio, silence_threshold)
 
+    # EXTENSION
+    # the different seed generation had to be implemented -> also removing the one hot encoding and the mu-quantization
+    # and reading hdf5 files instead of wav files
     #quantized = mu_law_encode(audio, quantization_channels)
 
 
@@ -258,6 +263,8 @@ def main():
 
         if (args.pca_out_path and args.save_every and (step + 1) % args.save_every == 0):
             #out = sess.run(decode, feed_dict={samples: waveform})
+            # EXTENSION
+            # does not need to decode the mu-quantization
             write_pca(waveform, wavenet_params['sample_rate'], args.pca_out_path)
 
     # Introduce a newline to clear the carriage return from the progress.
@@ -278,6 +285,8 @@ def main():
         write_wav(out, wavenet_params['sample_rate'], args.wav_out_path)
 
     if args.pca_out_path:
+        # EXTENSION
+        # remove the decoding of the one-hot-vector and remove also the mu-quantization.
         #out = sess.run(decode, feed_dict={samples: waveform})
         write_pca(waveform, wavenet_params['sample_rate'], args.pca_out_path)
 

### 2017-01-20 normalize.py

@@ -1,43 +1,36 @@
-import librosa
 import numpy as np
-import matplotlib.pyplot as plt
-from scipy import fft
 import intervaltree
-from sklearn.decomposition import IncrementalPCA
 import h5py
 import argparse
 
-fs = 44100
-crop_freq_th = 150
-window_size = 2048
-window_size = 2048  # 2048-sample fourier windows
-stride = 512        # 512 samples between windows
 
-coeff_per_window = 100
-
-silence_th = 0.2
-
-wps = fs/float(stride) # ~86 windows/second (with a fs of 16000)
-
-VERBOSE = True
-
+'''
+The function normalize takes as input the pca 
+coefficients from the file 'pca_file'
+and normalizes them to 0 mean and unit variance.
+The resulting coefficients are stored in output_file.
+'''
 def normalize(pca_file, output_file):
-    if not output_file:
-        output_file = pca_file[:-3] + "_pca.h5"
 
+    # open original pca file
     pca_input = h5py.File(pca_file, 'r')
     pca_input = pca_input['coeff']
 
     pca_data = []
+
+    # for each sample: add the coefficients
     for key in pca_input:
         pca_data += pca_input[key]
 
-    #pca_input = h5py.File(pca_file, 'r')['coeff'].value
+    # compute the mean and the variance
     mean = np.mean(pca_data)
     var = np.var(pca_data)
+
+    # subtract the mean and the variance
     pca_data -= mean
     pca_data /= var
 
+    # store the normalized values again
     h5f = h5py.File(output_file, 'w')
     h5f.create_dataset('normalization/mean_', data=mean)
     h5f.create_dataset('normalization/var_', data=var)
@@ -52,5 +45,3 @@ if __name__ == '__main__':
 	args = parser.parse_args()
 
 	normalize(args.data_file, args.output_file)
-
-	#np.save(output_file, proc_data)

### 2017-01-20 postprocess.py

@@ -1,0 +1,34 @@
+import preprocess
+import h5py
+import argparse
+import numpy as np
+import librosa
+
+sr = 44100
+
+def write_wav(waveform, sample_rate, filename):
+	'''
+	Save waveform as wav file
+	'''
+	y = np.array(waveform)
+	librosa.output.write_wav(filename, y, sample_rate)
+	print('WAV written at {}'.format(filename))
+
+def postprocess(pca_file, coeff_file, wav_file):
+
+	# load the frequencies of the produced file
+	freq = preprocess.load_pca(pca_file, coeff_file)
+
+	# convert the frequencies in 1d audio signal
+	audio = preprocess.load_freq(freq)
+
+	write_wav(audio, sr, wav_file)
+
+if __name__ == '__main__':
+	parser = argparse.ArgumentParser()
+	parser.add_argument("--pca_file", type=str, required=True, help="Path of the pca mean / variance input file")
+	parser.add_argument("--coeff_file", type=str, required=True, help="Path of the pca coefficient input file")
+	parser.add_argument("--wav_file", type=str, required=True, help="Path of the wav output")
+	args = parser.parse_args()
+
+	postprocess(args.pca_file, args.coeff_file, args.wav_file)

### 2017-01-20 preprocess.py

@@ -1,6 +1,5 @@
 import librosa
 import numpy as np
-import matplotlib.pyplot as plt
 from scipy import fft
 import intervaltree
 from sklearn.decomposition import IncrementalPCA
@@ -9,21 +8,21 @@ import argparse
 import os
 import csv
 
-fs = 44100
-crop_freq_th = 150
+crop_freq_th = 150 # from the originally 2048 frequencies only the first 150 are considered relevant
 window_size = 2048  # 2048-sample fourier windows
 stride = 512 #samples between windows
 
-coeff_per_window = 100
+coeff_per_window = 100 # the number of pca coefficients considered as relevant
 
-silence_th = 0.2
+silence_th = 0.2 # the value for which to trim the silence at the beginning and at the end of a piece
 
-wps = fs/float(stride) # ~86 windows/second (with a fs of 16000)
 
 VERBOSE = True
 
 def trim_silence(audio, threshold):
-	'''Removes silence at the beginning and end of a sample.'''
+	'''
+	Removes silence at the beginning and end of a sample.
+	'''
 	energy = librosa.feature.rmse(audio)
 	frames = np.nonzero(energy > threshold)
 	indices = librosa.core.frames_to_samples(frames)[1]
@@ -32,16 +31,29 @@ def trim_silence(audio, threshold):
 	return audio[indices[0]:indices[-1]] if indices.size else audio[0:0]
 
 def freq_extraction(data):
-	# coeffs/second = coeff_per_window * wps
+	'''
+	Requires the raw audio data (1d-array with amplitudes).
+	Returns a 2d array with the frequencies of the audio sample.
+	The returned frequencies are cropped to the first 'crop_freq_th' ones
+	(ignoring unimportant frequencies)
+	'''
+
+	# remove the silence at the beginning and at the end of the piece
 	data = trim_silence(data, silence_th)
 
 	data = data.copy()
 
+	# freq_length indicates to how many time steps the audio piece gets reduced 
+	# (i.e. how many fourier windows are generated)
 	freq_length = int(np.ceil((len(data)-window_size)/stride))+1
+
+	# the data is resized s.t. the audio fits exactly the window_size and the stride
 	data.resize(int((freq_length-1)*stride+window_size))
 
+	# initialize empty result variable
 	freq = np.empty([freq_length, crop_freq_th], dtype=complex)
 
+	# compute the fft of the fourier window with size window_size for every stride
[…]

### 2017-01-20 wavenet/audio_reader.py

@@ -40,29 +40,19 @@ def load_vctk_audio(directory, sample_rate):
         speaker_id, recording_id = [int(id_) for id_ in matches]
         yield audio, speaker_id
 
-def load_npz_audio(directory, sample_rate):
-    files = find_files(directory, pattern='*.npz')
-    for filename in files:
-        keys = []
-        data = np.load(open(filename, 'rb'), encoding='bytes')
-        if os.path.isfile(filename[:-4]+"_metadata.csv"):
-            with open(filename[:-4]+"_metadata.csv", 'r') as f:
-                reader = csv.reader(f)
-                for row in reader:
-                    if row[1] == "Beethoven" and row[2].find("Piano") >= 0:
-                        keys.append(row[0])
-        else:
-            keys = data.files
-        for file_i in keys:
-            X,Y = data[str(file_i)]
-            X = X.astype("float32")
-            X = X.reshape(-1,1)
-            yield X, '{}_{}'.format(filename, file_i)
+
+'''
+EXTENSION
+this loads the data in pca format to feed into the network.
+'''
 
 def load_pca_audio(directory, sample_rate):
+    # search for all h5 files
     files = find_files(directory, pattern='*.h5')
     for filename in files:
         h5f = h5py.File(filename, 'r')
+
+        # The coefficients are read and the (key (id), value (pca coefficients)) - pair is produced
         keys = h5f['coeff']
         for file_i in keys:
             X = h5f['coeff/{}'.format(file_i)].value
@@ -122,6 +112,9 @@ class AudioReader(object):
                 if self.coord.should_stop():
                     stop = True
                     break
+                # EXTENSION
+                # this function had to be adapted. The one-hot encoding had to be removed
+
 
                 # Silence threshold not needed with pca data, since this is alrady taken into account when producing the pca data
                 #if self.silence_threshold is not None:

### 2017-01-20 wavenet/model.py

@@ -425,7 +425,14 @@ class WaveNetModel(object):
         '''Computes the probability distribution of the next sample based on
         all samples in the input waveform.
         If you want to generate audio by feeding the output of the network back
-        as an input, see predict_proba_incremental for a faster alternative.'''
+        as an input, see predict_proba_incremental for a faster alternative.
+
+    
+        EXTENSION
+        remove the softmax layer
+        And no encoding as a one hot vector
+        -> The output of the network is taken as it is.
+        '''
         with tf.name_scope(name):
             if self.scalar_input:
                 encoded = tf.cast(waveform, tf.float32)
@@ -451,7 +458,14 @@ class WaveNetModel(object):
     def predict_proba_incremental(self, waveform, name='wavenet'):
         '''Computes the probability distribution of the next sample
         incrementally, based on a single sample and all previously passed
-        samples.'''
+        samples.
+
+
+        EXTENSION
+        remove the softmax layer
+        And no encoding as a one hot vector
+        -> The output of the network is taken as it is.
+        '''
         if self.filter_width > 2:
             raise NotImplementedError("Incremental generation does not "
                                       "support filter_width > 2.")
@@ -484,6 +498,11 @@ class WaveNetModel(object):
         '''Creates a WaveNet network and returns the autoencoding loss.
 
         The variables are all scoped to the given name.
+
+
+        EXTENSION
+        Change the loss function to the mean squared error loss of the resulting pca coefficients vectors (length 100)
+        The one hot decoding is removed and the cross-entropy loss too.
         '''
         with tf.name_scope(name):
 
@@ -495,7 +514,7 @@ class WaveNetModel(object):
 
             #encoded = self._one_hot(input_batch)
 
-            # New: no encoding of the input
+            # EXTENSION: no encoding of the input
             encoded = input_batch
             if self.scalar_input:
                 network_input = tf.reshape(
@@ -518,8 +537,9 @@ class WaveNetModel(object):
                 #loss = tf.nn.softmax_cross_entropy_with_logits(
                 #    prediction,
                 #    tf.reshape(shifted, [-1, self.quantization_channels]))
+
+                # EXTENSION: new loss function: mean squared error
                 loss = tf.square(prediction-tf.reshape(shifted, [-1, self.quantization_channels]))
-                #try out sum
                 reduced_loss = tf.reduce_mean(loss)
 
                 tf.scalar_summary('loss', reduced_loss)

### 2017-01-20 paper/paper.tex
added trainig error plot again.

@@
 \begin{abstract}
-  A critical part of scientific discovery is the
-  communication of research findings to peers or the general public.
-  Mastery of the process of scientific communication improves the
-  visibility and impact of research. While this guide is a necessary
-  tool for learning how to write in a manner suitable for publication
-  at a scientific venue, it is by no means sufficient, on its own, to
-  make its reader an accomplished writer. We also describe the rules
-  for submission in the computational intelligence laboratory.
-  This guide should be a
-  starting point for further development of writing skills.
+	This paper introduces a new attempt to generate audio by a deep neural network based on a stack of dilated convolutional layers.
+It adopts the network architecture from the paper "WaveNet: A Generative Model for Raw Audio" to operate in the frequency space.
+When applied to piano recording, the network is able to learn low basis frequencies, but struggles to grasp and generate sound features of dynamic and well-sounding audio.
+Nevertheless, different attempts of normalization and data preprocesing yield conceptual insights in the functioning of the WaveNet architecture.
 \end{abstract}
@@
 
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/loss.png}
+  \caption{Training error}
+  \label{fig:training_error}
+\end{figure}
 \section{Results}
-The results of all experiments are summarized in a python notebook of the GitHub repository.\footnote{"evaluate\_results.ipynb", available in the subfolder "notebooks"} Unfortunately, the adopted network could not perform as well as the original WaveNet architecture.
+The results of all experiments are summarized in a python notebook of the GitHub repository.\footnote{"evaluate\_results.ipynb", available in the subfolder "notebooks"} A suitable training session required approximately 6000-8000 training steps.
+Figure \ref{fig:training_error} plots the typical process of the error during training.\\
+ Unfortunately, the adopted network could not perform as well as the original WaveNet architecture.
 All generated audio files stabilized on one distribution of frequencies after a few seconds.
@@
 
-%\begin{figure*}[tbp]
-%  \centering
-%  \includegraphics[width=\textwidth]{figures/loss.png}
-%  \caption{Training error}
-%  \label{fig:training_error}
-%\end{figure*}
-
 \section{Discussion}

### 2017-01-20 paper/paper.tex
corrected caption of figure 4
added stride section to discussion

@@
   \includegraphics[width=\columnwidth]{figures/pca_explained_variance}
-  \caption{Signal compression and denoising using the Daubechies wavelet basis.}
+  \caption{Explained variance of the i'th eigenvector in the principal component analysis (with i being the horizontal axis)}
   \label{fig:pca_explained_variance}
@@
 Figure \ref{fig:generated_frequencies} visualizes the frequencies of the first 30 seconds of the output.
-In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.\textcolor{red}{Missing for discussion: on the one hand overlap necessary, on the other hand emphasises this to learn constant tones.}\\
+In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.\\
 A second conceptual attempt during the experiments was to normalize the PCA results before the training.
@@
 
+It has been shown that the stride is of great importance for the strength of the frequencies.
+Strong frequencies are interpreted as the network being confident in its prediction while weak frequencies are a sign of uncertainty.
+When the training timesteps do not overlap (stride = window size), the network produces much weaker frequencies.
+But a big stride might also lead to what we experience as not-changing frequencies.
+With a small stride most of the data chunks are the same as the previous one (when the length of a tone is much longer than one time unit).
+With larger strides, the frequencies change abruptly from one window to another.
+When a change in frequencies occurrs on an dataset with smaller strides, the absolute value of the difference of frequencies is much smaller.\\
+Therefore one can assume that with changing the setting (stride, but also window size, dilations, normalization etc.), also this approach might produce good results.
+
 \section{Summary}

### 2017-01-20 paper/paper.tex
refactored session "data preprocessing"

--- 2017-01-20 paper/paper.tex
@@
 As training data in the experiments, piano samples from the MusicNet dataset, a collection of 330 freely-licenced recordings of classical music are taken.
+The music is recorded with 44100 samples per second.
 The data is publicly available online \cite{musicnet}.\\
@@
 These 100 PCA coefficient are considered one tone and fed into the network as one time step.\\
-The eigenvectors of the PCA are shown in figure \ref{fig:pca_eigenvectors_imaginary_numbers}.
-Figure \ref{fig:pca_eigenvectors} illustrates how the PCA looks like when taking the absolute values of the imaginary numbers, out of which the components are built from.
-In this figure, one can see, that the relevant frequencies are all relatively low.
-Figure \ref{fig:pca_explained_variance} shows that a vast majority of the variance of the frequencies is captured by the first 100 PCA components.
-Summarizing, in neither of the two steps, we loose much of the original input information.
-This should lead to a decent compression of the audio signal.\\
-To make this evident, one can figure invert all the steps and plot the compressed wave and the original wave (see \ref{fig:preprocessing_difference_waveform}).
-It is easy to see that they resemble a lot.
-After preprocessing the audio contains a cracking noise, but nevertheless is the melody and the tone of the music well audible.\\
-To ensure this preprocessing still preserves the nature of the music, one can invert all steps.
-Although after preprocessing, there is an audible cracking, the music then is still clearly hearable and the pitches recognizable.
-On one hand, the waveform looks still similar (figure \ref{fig:preprocessing_difference_waveform})
-Whereas in the original WaveNet architecture, 16000 samples represented on second of audio, the adopted network only needs to process $16000/2048 \approx 8$ samples per second.
+Whereas in the original WaveNet architecture, 16000 samples represented on second of audio, the adopted network only needs to process $44100/2048 \approx 22$ samples per second.
 However, this change infers a fundamentally different approach of audio training and generation, described in the next section.
+\\
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\columnwidth]{figures/pca_eigenvectors_imaginary_numbers.png}
-  \caption{Eigenvectors of the PCA where the imaginary and the real part of the FFT are stored as a separate value in the PCA component.}
-  \label{fig:pca_eigenvectors_imaginary_numbers}
-\end{figure}
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\columnwidth]{figures/pca_eigenvectors.png}
-  \caption{Transformed eigenvectors of the PCA: The values of the vectors are the absolute value of the corresponding imaginary number}
-  \label{fig:pca_eigenvectors}
-\end{figure}
+\subsection{Evaluation of the preprocessing}
+The following analyses the amount of information which is lost by the compression steps of the previous session.
+Figure \ref{fig:pca_explained_variance} plots the explained variance of the first 250 eigenvectors during the The vast majority of the variance of the frequencies is captured by the first 100 PCA components.
+Thus, PCA is a reasonable measure to reduce the 300 dimensional data from the FFT to 100 PCA components.\\
+The best way to evaluate a compression procedure is to compare a compressed and afterwards reconstructed input to the original data.
+Figure \ref{fig:preprocessing_difference_waveform} compares exemplarily the recorded raw audio input of a music sample from the MusicNet dataset to a compressed and afterwards reconstructed version of the signal.
+The plot shows that the two signals are similar in their general behavour and only vary vary only in small details.\\
+The same results can be observed by listening to the original and the compressed and reconstructed audio file: After preprocessing the audio contains some cracking noise, but nevertheless is the melody and the tone of the music well audible.\\
+Summarizing, in neither by cutting after the first 150 FFT coefficients, nor by the PCA sigifican information of the original input is lost.
+Therefore, the described preprocessing leads to a decent compression of the audio signal.\\
 
@@
   \includegraphics[width=\columnwidth]{figures/preprocessing_difference_waveform}
-  \caption{An example of how the output audio differs before and after preprocessing (compression)}
+  \caption{An example of how the output audio differs before and after compression.}
   \label{fig:preprocessing_difference_waveform}
@@
 Figure \ref{fig:training_error} plots the typical process of the error during training.\\
+\subsection{General results}
  Unfortunately, the adopted network could not perform as well as the original WaveNet architecture.
@@
 Nevertheless, the experiments of this paper can be analysed more detailled. Varying the stride and including normalization yielded conceptually different results.\\
+\subsection{Striding}
 Overlapping chunks of input data turned out to be essential for learning audio features properly.
@@
 In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.\textcolor{red}{Missing for discussion: on the one hand overlap necessary, on the other hand emphasises this to learn constant tones.}\\
+\subsection{Normalization}
 A second conceptual attempt during the experiments was to normalize the PCA results before the training.

### 2017-01-20 paper/paper.tex
removed comment

@@
 Figure \ref{fig:generated_frequencies} visualizes the frequencies of the first 30 seconds of the output.
-In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.\textcolor{red}{Missing for discussion: on the one hand overlap necessary, on the other hand emphasises this to learn constant tones.}\\
+In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.
 \subsection{Normalization}

### 2017-01-20 paper/paper.tex
refactored summary

@@
 \section{Summary}
-With WaveNet an already very promising music generation network is presented.
+With WaveNet a very promising music generation network is presented.
 It relies on predicting amplitudes when the preceeding amplitudes are given.
@@
 But at the same time, the ability to quantize the network's outputs is lost.
-Unfortunately the network returned results not as good as with the original WaveNet paper.
-With the new method, the generated audio files rapidly converge to one steady tone.
-The changing behaviour of pitches is not learned by the network.
+Unfortunately the network could not return as good results as the original WaveNet paper.
+With the adopted network architecture, the generated audio files rapidly converge to one steady tone, while changes of pitches is not learned.
+However, by the approaches and discussions in the paper yiels conceptual insights and motivate further investigations.
 

### 2017-01-20 paper/paper.tex
added pdf, fixed typos

@@
 The aim of this paper is to enhance the WaveNet architecture and apply it to a more feature-oriented representation of input data.
-Instead of performing a classification task on raw audio, the network is adopted to operate on the frequencies of the audio data, processed by Fouier transformation.\\
-As training data in the experiments, piano samples from the MusicNet dataset, a collection of 330 freely-licenced recordings of classical music are taken.
+Instead of performing a classification task on raw audio, the network is adopted to operate on the frequencies of the audio data, processed by Fourier transformation.\\
+As training data in the experiments, piano samples from the MusicNet dataset, a collection of 330 freely-licensed recordings of classical music are taken.
 The music is recorded with 44100 samples per second.
-The data is publicly available online \cite{musicnet}.\\
+The data is publicly available on the Internet \cite{musicnet}.\\
 In particular, the following topics are covered:
@@
 By using exponentially growing dilations in depth, a network with $n$ dilated convolutional layers can observe the last $2^n$ input values to predict an output.\\
-Training WaveNet is done similar as an autoencoder: The $i$th amplitude of the input data is predicted by only using previous input data and taken as label for the loss calculation and backpropagation.
+Training WaveNet is done similar as an auto-encoder: The $i$th amplitude of the input data is predicted by only using previous input data and taken as label for the loss calculation and backpropagation.
 Thus, training only requires the input data and can be done in parallel.
@@
 Figure \ref{fig:preprocessing_difference_waveform} compares exemplarily the recorded raw audio input of a music sample from the MusicNet dataset to a compressed and afterwards reconstructed version of the signal.
-The plot shows that the two signals are similar in their general behavour and only vary vary only in small details.\\
+The plot shows that the two signals are similar in their general behavior and only vary vary only in small details.\\
 The same results can be observed by listening to the original and the compressed and reconstructed audio file: After preprocessing the audio contains some cracking noise, but nevertheless is the melody and the tone of the music well audible.\\
-Summarizing, in neither by cutting after the first 150 FFT coefficients, nor by the PCA sigifican information of the original input is lost.
+Summarizing, in neither by cutting after the first 150 FFT coefficients, nor by the PCA significant information of the original input is lost.
 Therefore, the described preprocessing leads to a decent compression of the audio signal.\\
-
 \begin{figure}[htbp]
@@
 This shows that the network has difficulties to decide which frequencies are important and cannot produce as compact and compressed audio output as it is achieved by using FFT and PCA on recorded audio data.\\
-Nevertheless, the experiments of this paper can be analysed more detailled. Varying the stride and including normalization yielded conceptually different results.\\
+Nevertheless, the experiments of this paper can be analyzed more detailed. Varying the stride and including normalization yielded conceptually different results.\\
 \subsection{Striding}
@@
 The aim of this additional preprocessing task was to balance the contribution of all 100 PCA dimensions to the loss, hoping that all frequencies are learned more equally.
-In a first attempt, data normalzed to zero mean and unit variance turned out to produce too small loss for training.
+In a first attempt, data normalized to zero mean and unit variance turned out to produce too small loss for training.
 With vanishing gradients the network could not learn at all.
@@
 \end{figure}
-
 \section{Discussion}
@@
   \item The original paper quantizes the amplitudes and can then use a softmax layer as a last layer of the network.
-In this new approach, we need continous values (PCA coefficients).
+In this new approach, we need continuous values (PCA coefficients).
 Since the results are much worse than with WaveNet, the missing quantization (and connected to it the missing softmax layer) might be the reason behind this phenomenon.
@@
 With larger strides, the frequencies change abruptly from one window to another.
-When a change in frequencies occurrs on an dataset with smaller strides, the absolute value of the difference of frequencies is much smaller.\\
+When a change in frequencies occurs on an dataset with smaller strides, the absolute value of the difference of frequencies is much smaller.\\
 Therefore one can assume that with changing the setting (stride, but also window size, dilations, normalization etc.), also this approach might produce good results.
@@
 With WaveNet a very promising music generation network is presented.
-It relies on predicting amplitudes when the preceeding amplitudes are given.
+It relies on predicting amplitudes when the preceding amplitudes are given.
 In this paper it is investigated whether transforming the data to the Fourier space improves the predictions.
@@
 With the adopted network architecture, the generated audio files rapidly converge to one steady tone, while changes of pitches is not learned.
-However, by the approaches and discussions in the paper yiels conceptual insights and motivate further investigations.
+However, by the approaches and discussions in the paper yields conceptual insights and motivate further investigations.
 
